Models of gravitational lens candidates from Space Warps CFHTLS

### 2017-08-30 paper2.tex
about mass map diagnostic
replace total mass with lensing mass 2
bigger labels for all plots
several edits
working on time
overhaul

% Only include extra packages if you really need them. Common packages are:
%\usepackage[dvipdfmx]{graphicx}	% Including figure files
\usepackage{graphicx} % Including figure files
%\usepackage[skip=0pt]{subcaption}
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{pifont}	% Extra maths symbols
%\usepackage{bmpsize}  % PS still needs this for correct bounding boxes?

@@

\newcommand*{\rot}{\rotatebox{90}}
\newcommand*{\OK}{\ding{51}}
\newcommand*{\NO}{\ding{55}}
\newcommand*{\UK}{--}

\newcommand{\lenstitle}[1]{\noindent\textbf{#1} --}
\newcommand{\params}[3]{(\(\Theta_\text{E}:#1\), $\varepsilon:#2$, $\alpha_\varepsilon:#3$ )}

\newcommand{\SW}{Space~Warps}
\newcommand{\SpL}{SpaghettiLens}
\newcommand{\asw}[1]{ASW000#1}
\newcommand{\sw}[1]{SW~#1}
\newcommand{\model}[1]{SL model~#1}

@@

@@

\begin{abstract}
We report modelling follow-up of recently-discovered
gravitational-lens candidates in the Canada France Hawaii Telescope
Legacy Survey. Lens modelling was done by a small group of
specially-interested volunteers from the \SW\ citizen-science community
who originally found the candidate lenses.  Models are categorised
according to seven diagnostics indicating (a)~the image morphology and
how clear or indistinct it is, (b)~whether the mass map and synthetic
lensed image appear to be plausible, and (c)~how the lens-model mass
compares with compares with the abundance-matched halo mass derived
from photometry of the lensing galaxy.

The lens-model masses range from $\sim10^{11}M_\odot$ to
$>10^{13}M_\odot$.  Preliminary estimates of the stellar masses show
(with two outliers) a smaller spread in stellar mass: a factor of a
few below or above $\sim10^{11}M_\odot$.  That is, the stellar-mass
fraction declines sharply as total mass increases.  The highest-mass
system with a convincing model is J1434+522 (SW05).  The two low-mass
outliers are J0206-095 (SW19) and J2217+015 (SW42); if these two are
indeed lenses, they are probing an interesting regime of very low
star-formation efficiency.

Some improvements to the modelling software used (SpaghettiLens), and
discussion of strategies for scaling to future surveys with more and
frequent discoveries, are also included.
\end{abstract}

>>>>

% Only include extra packages if you really need them. Common packages are:
%\usepackage[dvipdfmx]{graphicx}	% Including figure files
\usepackage{graphicx} % Including figure files
%\usepackage[skip=0pt]{subcaption}
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{pifont}	% Extra maths symbols
%\usepackage{bmpsize}  % PS still needs this for correct bounding boxes?
\usepackage{xspace} % handling of spaces after shortcuts

@@

\newcommand*{\rot}{\rotatebox{90}}
\newcommand*{\OK}{\ding{51}}
\newcommand*{\NO}{\ding{55}}
\newcommand*{\UK}{\ensuremath{\text{--}}}

%\newcommand{\lenstitle}[1]{\noindent\textbf{#1} --}
%\newcommand{\params}[3]{(\(\Theta_\text{E}:#1\), $\varepsilon:#2$, $\alpha_\varepsilon:#3$ )}

\newcommand{\SW}{Space\,Warps\xspace}
\newcommand{\SpL}{SpaghettiLens\xspace}
\newcommand{\asw}[1]{ASW000#1\xspace}
\newcommand{\sw}[1]{SW\,#1\xspace}
%\newcommand{\model}[1]{SL model~#1}

@@

\newcommand{\sqdeg}{\ensuremath{\text{deg}^2}}

@@

\begin{abstract}
We report modelling follow-up of recently-discovered
gravitational-lens candidates in the Canada France Hawaii Telescope
Legacy Survey. Lens modelling was done by a small group of
specially-interested volunteers from the \SW\ citizen-science community
who originally found the candidate lenses.  Models are categorised
according to seven diagnostics indicating (a)~the image morphology and
how clear or indistinct it is, (b)~whether the mass map and synthetic
lensed image appear to be plausible, and (c)~how the lens-model mass
compares with compares with the abundance-matched halo mass derived
from photometry of the lensing galaxy.
The lens-model masses range from $\sim10^{11}M_\odot$ to
$>10^{13}M_\odot$.  Preliminary estimates of the stellar masses show
(with two outliers) a smaller spread in stellar mass: a factor of a
few below or above $\sim10^{11}M_\odot$.  That is, the stellar-mass
fraction declines sharply as lensing mass increases.  The highest-mass
system with a convincing model is J1434+522 (\sw{05}).  The two low-mass
outliers are J0206-095 (\sw{19}) and J2217+015 (\sw{42}); if these two are
indeed lenses, they are probing an interesting regime of very low
star-formation efficiency.
Some improvements to the modelling software used (SpaghettiLens), and
discussion of strategies for scaling to future surveys with more and
frequent discoveries, are also included.
\end{abstract}

### 2017-08-30 parts/figures.tex
big table update
overhaul

\begin{figure*}
\includegraphics[width=\linewidth]{img/mlens_vs_mstel_one/mstel_vs_mtot_one}
\caption{Total mass in the model against the estimated stellar mass,
  alongside the values for the whole sample.  (The labelled orange
  bars are the systems shown in detail in
  Figures~\ref{fig:splinput}--\ref{fig:encl}.)  The horizontal extent
  of each bar indicates the extreme cases of a young (0.5~Gyr-old)
  stellar population and the oldest possible population at the given
  redshift.  The vertical extent indicates the spread of masses in
  lens-model ensemble.  The lower-right shaded region is unphysical
  according to the stellar-population models, because it gives
  $M<\Mstel$. The upper-left shaded region is unphysical according to
  abundance matching (see Section~\ref{sec:stellar-mass}) because it
  gives $M>\Mhalo$.  That is to say, the unshaded region is
  $0<\haloindex<1$. \label{fig:stelmass}}
\end{figure*}

\begin{table*}
  \caption{Diagnostics of a selected model for each Space Warps candidate
    (see Section~\ref{sec:summary}).}
  \label{tab:models}
  \include{parts/table}
\end{table*}

>>>>

\begin{figure*}
\includegraphics[width=\linewidth]{img/mlens_vs_mstel_one/mstel_vs_mtot_one}
\caption{Total mass in the model against the estimated stellar mass,
  alongside the values for the whole sample.  (The labelled orange
  bars are the systems shown in detail in
  Figures~\ref{fig:splinput}--\ref{fig:encl}.)  The horizontal extent
  of each bar indicates the extreme cases of a young (0.5~Gyr-old)
  stellar population and the oldest possible population at the given
  redshift.  The vertical extent indicates the spread of masses in the
  lens-model ensemble.  The lower-right shaded region is unphysical
  according to the stellar-population models, because it gives
  $M<\Mstel$. The upper-left shaded region is unphysical according to
  abundance matching (see Section~\ref{sec:stellar-mass}) because it
  gives $M>\Mhalo$.  That is to say, the unshaded region is
  $0<\haloindex<1$. \label{fig:stelmass}}
\end{figure*}

\begin{table*}
  \caption{Diagnostics of a selected model for each Space Warps candidate
    (see Section~\ref{sec:summary}). Machine readable version in online supplement.}
  \label{tab:models}
  \include{parts/table}
\end{table*}

### 2017-08-30 parts/intro.tex
overhaul

This paper continues that study by applying SpaghettiLens to
lens-candidates discovered through Space~Warps.  We present results
from modelling of 56 of the 59 lens candidates reported by
\cite{2016MNRAS.455.1191M}.  Each lens candidate was modelled in a
collaborative refinement process, where anyone interested can create a
new model or modify an existing model to try and make it
better.\footnote{This is in contrast to the main Space~Warps project
  for discovering lenses, where volunteers in a crowd of $\gtrsim10^4$
  make independent contributions.  Each person is presented with a
  random selection of survey-patches and invited to (in effect) vote
  on each.  The system estimates each volunteer's skill level
  according to test-patches interspersed with the real data, and
  weights their votes accordingly \citep{2016MNRAS.455.1171M}.  There
  is an active forum for volunteers, but since everyone is seeing
  different data samples with minimal overlap, the forum has little if
  any influence on votes.}  The resulting model represents a consensus
among contributors, as to the best that could be achieved with the
available data and software.

@@

TODO: very short summary of the following sections

@@

\section{The candidates and models}

TODO: About SW sample

TODO: about SpL

@@

The volunteers were encouraged to create models of objects found in \SW before they were classified as plausible candidates.
Figure~\ref{fig:time} shows the total number of models generated over time, models created for systems that turned out to be non candidates were not included.
The first beta version of \SpL was launched around day 140 and additional features were implemented on the fly.
The system was tested by eight major contributors, volunteers that at least generated five models.
Around day 200 (Aug. 2013) a modelling challenge for the candidate later to be identified as \sw{01} was launched.
A group of five volunteers converged to a set of favorite models that they presented in an online letter after about 65 days and 30 models have been created for this particular system.
Around day 820 (Apr. 2015) the first full version of \SpL was presented for testing, around the same time the list of identified candidates was available \citep[as a preprint of the later work][]{2016MNRAS.455.1191M}.
At this stage, after a call to model candidates that have been neglected up to that date, candidate \sw{29} shows that the volunteers converged to a favorable model after 15 models generated in 30 days.

>>>>

This paper continues that study by applying SpaghettiLens to
lens-candidates discovered through Space~Warps.  We present results
from modelling of 58 of the 59 lens candidates reported by
\cite{2016MNRAS.455.1191M}.  Each lens candidate was modelled in a
collaborative refinement process, where anyone interested can create a
new model or modify an existing model to try and make it
better.\footnote{This is in contrast to the main Space~Warps project
  for discovering lenses, where volunteers in a crowd of $\gtrsim10^4$
  make independent contributions.  Each person is presented with a
  random selection of survey-patches and invited to (in effect) vote
  on each.  The system estimates each volunteer's skill level
  according to test-patches interspersed with the real data, and
  weights their votes accordingly \citep{2016MNRAS.455.1171M}.  There
  is an active forum for volunteers, but since everyone is seeing
  different data samples with minimal overlap, the forum has little if
  any influence on votes.}  The resulting model represents a consensus
among contributors, as to the best that could be achieved with the
available data and software.

@@

This paper is organized as follows:
Section~\ref{sec:candidates_models} introduces the candidates, the models and the diagnostics applied to them.
The following sections elaborate on the diagnostics.
Image morphology diagnostics are explained in Section~\ref{sec:morph} and
diagnostics based on the mass models in Section~\ref{sec:massmodels}.
Stellar masses are calculated in Section~\ref{sec:stellar-mass} to compare stellar to lensing masses of the lensing candidates.
Section~\ref{sec:summary} gives our summary and conclusions.

@@

\section{The candidates and models}
\label{sec:candidates_models}

\SW is a citizen science gravitational lens search \citep{2016MNRAS.455.1171M}.
Its first run searched the CFHT Legacy Survey, a survey carried out in five optical bands ($u^*$,$g'$,$r'$,$i'$,$z'$) covering a total area of 160 \sqdeg, divided up into four fields.
It has a field of view of 1 \sqdeg, a pixel size of 0.186'' and the deepest limiting magnitude is 25.47 in the $g$ band\citep{2012AJ....143...38G}.
The cutouts shown to the volunteers were colour composites of the $g$, $r$ and $i$ band of randomly chosen regions.
It resulted in 59 new (of which 29 promising) lens candidates that have not been found by other searches already and in the rediscovery of 82 previously known lens candidates.
The candidates have broadly similar redshifts (0.2 -- 1.0), Einstein radii (0.7\arcsec -- 5\arcsec) and arc magnitudes (22 -- 26) as previous robotic robotic searches by ArcFinder and RingFinder \citep{2016MNRAS.455.1191M}.

@@

While \SW was running, volunteers started working with \SpL to create models of possible lensing candidates found in \SW that were debated on the \SW discussion forums, before they were classified as plausible candidates.
Figure~\ref{fig:time} shows the total number of models generated over time, models created for systems that turned out to be non candidates were not included.
The first beta version of \SpL was launched in May 2013 and additional features were implemented on the fly.
The system was tested by eight major contributors, volunteers that at least generated five models.
In the beginning of August 2013 (day 70) a modelling challenge for the candidate later to be identified as \sw{01} was launched.
A group of five volunteers converged to a set of favourite models that they presented in an online letter after about 65 days and 30 models have been created for this particular system.
In April 2015 (day 700) the first full version of \SpL was presented for testing.
In the same time the list of identified candidates was made available \citep[as a preprint of the later work][]{2016MNRAS.455.1191M}.
At this stage, the volunteers were asked to complete the sample and to create models for all the lens candidates that were still missing.
The generated models for candidate \sw{29} show that the volunteers converged to a favorable model much faster, after 15 models generated in 30 days.
This modelling effort resulted in a set of 356 \SpL models for all but one of the \SW 59 lens candidates; candidate \sw{39} was not modelled.
